title: "p8105_hw5_hy2912"
author: "Helen Yousaf"
date: "2024-11-05"
output: html_document
---
# Problem 2 
```{r}
library(tidyverse)
library(rvest)
library(broom)
library(ggplot2)
library(dplyr)
```

## Setting parameters 

```{r}
n <- 30
sigma <- 5
mu <- 0
alpha <- 0.05 
num_datasets <- 5000
```

## Vectors 
```{r}
mu_hat <- numeric(num_datasets)
p_values <- numeric(num_datasets)
```

## Simulation 
```{r}
set.seed(123) # For reproducibility 
for (i in 1:num_datasets) {
  # Generate data 
 x <- rnorm(30, mean = 0, sd = 5)
 
 # Conduct t-test 
 t_test_results <- t.test (x, mu=0)
 
 # Estimate and p-value 
  tidy_result <- tidy(t_test_results)
  mu_hat[i] <- tidy_result$estimate
  p_values[i] <- tidy_result$p.value
}
```

## Repeating for various mu values
```{r}
set.seed(12)
n <- 30
sigma <- 5
mu_values <- 1:6
num_datasets <- 5000
alpha <- 0.05
```

```{r}
power_results <- data.frame(mu = mu_values, power = numeric(length(mu_values)), avg_estimate = numeric(length(mu_values)), avg_estimate_rejected = numeric(length(mu_values)))

for (i in seq_along(mu_values)) {
  mu <- mu_values[i]
  p_values <- numeric(num_datasets)
  estimates <- numeric(num_datasets)
  
  for (j in 1:num_datasets) {
    x <- rnorm(n, mean = mu, sd = sigma)
    test_result <- t.test(x, mu = 0)
    tidy_result <- tidy(test_result)
    p_values[j] <- tidy_result$p.value
    estimates[j] <- tidy_result$estimate
  }
  
  power_results$power[i] <- mean(p_values < alpha)
  power_results$avg_estimate[i] <- mean(estimates)
  power_results$avg_estimate_rejected[i] <- mean(estimates[p_values < alpha])
}
```

```{r}
ggplot(power_results, aes(x = mu)) +
  geom_line(aes(y = power), color = "blue") +
  labs(y = "Power", x = "True Value of Î¼") +
  ggtitle("Power of One-Sample t-Test")

ggplot(power_results, aes(x = mu)) +
  geom_line(aes(y = avg_estimate), color = "red") +
  geom_line(aes(y = avg_estimate_rejected), color = "green") +
  labs(y = "Average Estimate of Î¼Ì‚", x = "True Value of Î¼") +
  ggtitle("Average Estimate of Î¼Ì‚ vs True Value of Î¼") +
  scale_color_manual(values = c("red", "green"), labels = c("Avg Estimate", "Avg Estimate (Null Rejected)")) +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = "Legend"))
```

This data is consistent with statistical theory that explains how effect size and power are directly proportional. Simply, as effect size increases, so does the power. In addition, the sample average of ðœ‡Ì‚ across tests for which the null is rejected is  approximately equal to the true value of ðœ‡ eventually when the graph equals out and as the true value of ðœ‡ increases. 

# Problem 3 

The raw data relies on information collected by the Washington Post on the more than 52,000 criminal homicides that have occurred in the past decade in all states across the largest US cities. The data captures information on: the location of the killing; whether an arrest was made and; and basic demographic information about each victim.

```{r}
library(purrr)
```

## Loading dataset
```{r}
url <- "https://raw.githubusercontent.com/washingtonpost/data-homicides/refs/heads/master/homicide-data.csv"
data <- read.csv(url)
```

## Creating a city_state variable
```{r}
data <- data %>%
  mutate(city_state = paste(city, state, sep = ", "))
```

## Summarizing homicides
```{r}
summary_data <- data %>%
  group_by(city_state) %>%
  summarize(total_homicides = n(),
            unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest")))
```

## Estimating proportion of unsolved homicides for Baltimore, MD
```{r}
baltimore_result <- prop.test(summary_data$unsolved_homicides[summary_data$city_state == "Baltimore, MD"],
                               summary_data$total_homicides[summary_data$city_state == "Baltimore, MD"])

baltimore_tidy <- broom::tidy(baltimore_result) %>%
  select(estimate, conf.low, conf.high)
```

## Running prop.test for each city and extract proportions and CIs

```{r}
results <- summary_data %>%
  rowwise() %>%
  mutate(prop_test = list(prop.test(unsolved_homicides, total_homicides))) %>%
  mutate(tidy_result = list(broom::tidy(prop_test))) %>%
  unnest(tidy_result) %>%
  select(city_state, estimate, conf.low, conf.high)
```

## Create a plot
```{r}
ggplot(results, aes(x = reorder(city_state, estimate), y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  labs(x = "City, State", y = "Proportion of Unsolved Homicides") +
  theme_minimal() +
  coord_flip()
```

